---
title: "3.2_Figures"
author: "Florian van Leeuwen"
date: "2023-02-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First results simulation study
The following parameters were used for the simulation study:
* Number of person: 100, 300, 500, 700, 900, 1200
* Number of time points: 6, 8, 10, 12, 14, 16, 18
* Number of simulations: 100
* Intercept: 170
* Variance of the intercept: N(0,25)
* Slope: 12.7
* Variance added after data generation: N(0,20)
* Effect sizes as % of the slope: 0.01, 0.02, 0.03, 0.04, 0.05, 0.06

The simulation was run for three conditions, see Figure below:
* Only a step change
* Only a slope change
* A step and slope change

```{r}
# figure of the sim studie
time <- rep(seq(10),3)
group <- c(rep("Step", 10), rep("Slope", 10), rep("Step and slope", 10))
score1 <- time * 2 + c(rep(0,5), rep(-2,5), rep(0,5), -0.5, -1.2, -2.2, -3.2, -4.2, rep(0, 5), 
                       -3, -4.5, -6, -7.5, -9)

data_eex <- data.frame(time = time, 
            group = group,
            score = score1)

data_eex %>% 
  ggplot(aes(x = time, y = score)) +
  geom_point() +
  geom_segment(aes(x = 1, y = 2, xend = 5.5, yend = 11, color = "General trend")) +
  geom_segment(aes(x = 5.5, y = 9, xend = 10, yend = 18, color = "Step")) +
  geom_segment(aes(x = 5.5, y = 11, xend = 10, yend = 16, color = "Slope")) +
  geom_segment(aes(x = 5.5, y = 8.75, xend = 10, yend = 11, color = "Step and slope")) +
  scale_color_discrete(name="Changing parameter") +
  labs(title = "The different simulation scenarios", y = "Score", x = "Time") +
  geom_vline(xintercept = 5.5) +
  scale_x_continuous("Time", labels=c("T_0", "(N+1)/2", "T_N"), breaks=c(1, 5.5, 10))

#ggsave("figures/sim_setup.png")
```

Each took about 30 minutes to compute (1.5 hour in total) using the following OLS model:

$$
Y_i = \beta_0 + \beta_1*Time + \beta_2*Intervention + \beta_3*Time*Interventios
$$

TOTAL NUMBER OF CONDITIONS: 6*7*6*3 = 756

The output of the simulation was assessed on three criteria:
* The average power
* The average bias
* The precision (number of times bias was > 5%)

```{r}
# import the data
out10 <- read.csv("results/sim10.csv")
out01 <- read.csv("results/sim01.csv")
out11 <- read.csv("results/sim11.csv")
```

### Scenario 1: Only a step change
```{r}
out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = pre_slope)) +
  geom_tile() + 
  labs(title = "Power for pre slope",x = "Number of time points",
       y = "Number of participants", fill="Power") +
  facet_wrap(~effect_sizes) +
  theme_minimal() 

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = step)) +
  geom_tile() + 
  labs(title = "Power for step",x = "Number of time points",
       y = "Number of participants", fill="Power") +
  facet_wrap(~effect_sizes) +
  theme_minimal() 

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = slope)) +
  geom_tile() + 
  labs(title = "Power for slope",x = "Number of time points",
       y = "Number of participants", fill="Power") +
  facet_wrap(~effect_sizes) +
  theme_minimal() 
```

```{r}
# heat maps
out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = bias_pre_slope)) +
  geom_tile() + 
  labs(title = "Percentage bias in pre slope", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = bias_step)) +
  geom_tile() + 
  labs(title = "Percentage bias in step", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = bias_slope)) +
  geom_tile() + 
  labs(title = "Percentage bias in slope", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()
```

```{r}
out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = Precision_pre_slope)) +
  geom_tile() + 
  labs(title = "Precision in pre slope in nsim = 1000", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = Precision_step)) +
  geom_tile() + 
  labs(title = "Precision in step in nsim = 1000", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()

out %>% 
  ggplot(aes(x = N_t, y = N_pers, fill = Precision_slope)) +
  geom_tile() + 
  labs(title = "Precisio in slope in nsim = 1000", x = "Number of time points", 
       y = "Number of participants", fill="Percentage bias") +
  facet_wrap(~effect_sizes) +
    theme_minimal()
```

### Scenario 2: Only a slope change
### Scenario 3: A step and slope change




#### another option for the simulation plot
```{r}
# figure of the sim studie
time <- rep(seq(10),3)
group <- c(rep("Step", 10), rep("Slope", 10), rep("Step and slope", 10))
score1 <- time * 2 + c(rep(0,5), rep(-2,5), rep(0,5), -0.5, -1.2, -2.2, -3.2, -4.2, rep(0, 5), 
                       -3, -4.5, -6, -7.5, -9)

data_eex <- data.frame(time = time, 
            group = group,
            score = score1)

P1 <- data_eex %>% 
  filter(group == "Step") %>% 
  ggplot(aes(x = time, y = score)) +
  geom_point() +
  geom_segment(aes(x = 1, y = 2, xend = 5.5, yend = 11)) +
  geom_segment(aes(x = 5.5, y = 9, xend = 10, yend = 18), color = "red") +
  guides(color="none") +
  ylim(0, 20) +
  labs(title = "Step", y = "Score", x = "Time") +
  geom_vline(xintercept = 5.5) +
  scale_x_continuous("Time", labels=c("T_0", "(N+1)/2", "T_N"), breaks=c(1, 5.5, 10)) +
  theme_classic()

P2 <- data_eex %>% 
  filter(group == "Slope") %>% 
  ggplot(aes(x = time, y = score)) +
  geom_point() +
  geom_segment(aes(x = 1, y = 2, xend = 5.5, yend = 11)) +
  geom_segment(aes(x = 5.5, y = 11, xend = 10, yend = 16), color = "green") +
  guides(color="none") +
  ylim(0, 20) +
  labs(title = "Slope", y = "Score", x = "Time") +
  geom_vline(xintercept = 5.5) +
  scale_x_continuous("Time", labels=c("T_0", "(N+1)/2", "T_N"), breaks=c(1, 5.5, 10)) +
  theme_classic()

P3 <- data_eex %>% 
  filter(group == "Step and slope") %>% 
  ggplot(aes(x = time, y = score)) +
  geom_point() +
  geom_segment(aes(x = 1, y = 2, xend = 5.5, yend = 11)) +
  geom_segment(aes(x = 5.5, y = 8.75, xend = 10, yend = 11),  color = "blue") +
  guides(color="none") +
  ylim(0, 20) +
  labs(title = "Step and slope", y = "Score", x = "Time") +
  geom_vline(xintercept = 5.5) +
  scale_x_continuous("Time", labels=c("T_0", "(N+1)/2", "T_N"), breaks=c(1, 5.5, 10)) +
  theme_classic()

P1 + P2 + P3

ggsave("figures/sim_setup1.png")
```
