---
title: "Sim_study_output"
author: "Florian van Leeuwen"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F, warning =F}
library(haven)
library(dplyr)
library(xtable)
library(simstudy)
library(lme4)
library(ggplot2)
library(tidyr)
library(patchwork)
library(lavaan)
```

```{r}
# Own functions
source("functions/1.Generate_data.R")
source("functions/2.Models.R")
source("functions/3.sim_all.R")
```

## TEST THE FUNCTIONS
```{r}
X <- forms_for_sim(N_time = 4, var = 10, Intercept = 170,  Slope = 16.5, step_size = 20, slope_size =  -10, additional_step = 0)
X
```

```{r, warning = F}
test <- gen_data(X$form, X$v_name, X$var, N_person = 100, X$Intercept, var_time = 0, additional_step_corr = 0)

# inspect data
test$long %>% 
  ggplot(aes(y = value, x = time, color = as.factor(id))) +
  geom_point(alpha = .7, size = .4) +
  geom_smooth(se = F, size = .5) +
  guides(color = FALSE)
```

```{r}
# check function for all Effect tizes
ES <- c(0.01, 0.025, 0.05, 0.10,0.15, 0.30)
All_conditions(ES, 12.7)
All_conditions(ES, 33)
```

## PARAMETERS SIMULATION STUDY
```{r}
# define input
N_person <- c(100, 300, 500, 700, 900, 1100, 1300, 1500)
N_time <- c(6, 8, 10, 12, 14, 16, 18, 20)
N_sim <- 1000
var <- 25
Intercept <- 170
Slope <- 16.5
var_time <- 25
ES <- c(0.01, 0.025, 0.05, 0.10, 0.15, 0.30)
```

# RUN THE DIFFERENT SCENARIO'S

## Step and slope change (section 5.1)
```{r}
set.seed(123)
out11 <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = -1, additional_step = 0, additional_step_corr = 0, time_2 = 0, model = "OLS")

write.csv(out11, "Output/results/sim11_final.csv")
```


## Only step change (section 5.2)
```{r}
set.seed(123)
# Basic setup
startTime <- Sys.time()
out10 <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = 0,additional_step = 0, additional_step_corr = 0, time_2 = 0, model = "OLS")
endTime <- Sys.time()

print(endTime - startTime)
write.csv(out10, "Output/results/sim10_final.csv")
```

## Only slope change (section 5.2)
```{r}
set.seed(123)
out01 <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = 0, treatment_slope = -1, additional_step = 0, additional_step_corr = 0, time_2 = 0, model = "OLS")

write.csv(out01, "Output/results/sim01_final.csv")
```

## Step and slope change and additional step change (section 5.3)
```{r}
set.seed(123)
out11_additional_step <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = -1, additional_step = 1, additional_step_corr = 0, time_2 = 0, model = "OLS")

write.csv(out11_additional_step, "Output/results/out11_additional_step_final.csv")
```

## Step and slope change and additional step change with correction (section 5.4)
```{r}
set.seed(123)
out11_additional_step_corr <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = -1, additional_step = 1, additional_step_corr = 1, time_2 = 0,  model = "OLS")

# write away the results
write.csv(out11_additional_step_corr, "Output/results/out11_additional_step_corr_final.csv")
```

## Step and slope change with correction for a changing slope (section 5.5)
```{r}
# correcting for time^2 even tough it is not in the data
time_2 <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = -1,additional_step = 0, additional_step_corr = 0, time_2 = 1,  model = "OLS")

write.csv(time_2, "Output/results/time2_final.csv")
```

### EXTRA 
```{r}
set.seed(123)
# different effect sizes 
out11_diff_effect <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -2, treatment_slope = -1, additional_step = 0, additional_step_corr = 0, time_2 = 0, model = "OLS")
```

```{r}
set.seed(123)
# One negative parameter
out11_neg_step <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = -1, treatment_slope = 1,additional_step = 0, additional_step_corr = 0, time_2 = 0,  model = "OLS")

write.csv(out11_neg_step, "Output/results/out11_neg_step_3.csv")
```

```{r}
set.seed(123)

out11_neg_slope <- sim_all(N_person, N_time, N_sim, var, Intercept, Slope, seed, var_time, ES, treatment_step = 1, treatment_slope = -1,additional_step = 0, additional_step_corr = 0, time_2 = 0, model = "OLS")
# write away the results
write.csv(out11_neg_slope, "Output/results/out11_neg_slope_3.csv")
```
